---
title: "Simulations_ToyModel"
author: "B. Raynor"
date: "3/27/2023"
output: html_document
---

## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path="figure")
```

Load packages, functions, data

```{r, message=FALSE, warning=FALSE}
#clear environment
rm(list = ls()) 

#load required packages
library(ggplot2)  #pretty figures
library(dplyr)    #tidy code
library(ggpubr)   #plot arrangement

source(here::here("R", "Functions_StochasticModels.R"))
source(here::here("R", "Functions_StochasticSimulationVisualization.R"))

#parameters
parms <- data.frame( 
  patch = c("Patch 1", "Patch 2", "Patch 3"),
  b= c(10,10,10),
  beta = c(0.001, 0.001, 0.001),
  mu = c(1/5000, 1/5000, 1/5000),
  alpha = c(1/3, 1/3, 1/3),
  nu1 = c(0,0,0),
  gamma = c(1/2, 1/2, 1/2),
  N = c(1000, 1000, 1000),
  current_infect = c(1,1,1))

knitr::kable(parms %>% select(-current_infect))

```

**Table:** Synthetic parameters used to explore general SEIV model

Initialize SDEs
```{r}
num_patch = length(parms$patch)

#initial conditions 
init <- data.frame(S=0.5*parms$N - 10*parms$current_infect,
                   E= rep(0,num_patch),
                   I= 10*parms$current_infect,
                   V= 0.5*parms$N)

#time series
num_years = 3
Time = 365*num_years #End time
dt = 1 #Step size dt

#How many simulations we want to average
sim=100

#pulse vax
vax0= matrix(0, Time/dt, num_patch)

```

## PART A: 3 patches all connected

```{r}
#contact matrix
df.contact <-data.frame(patch1 = c(0,1,1), patch2= c(1,0,1), patch3 =c(1,1,0))
beta.contact_constant = 0.00001 
beta.contact = rfun.FormatContact(df.contact, beta.contact_constant)

```

Sim 1: Single pulse, all vaccinated

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)),] = c(rep(0.7, num_patch)) #

#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

#assess time to elimination

df.system_elimination <- data.frame(avg_time_elimination = NULL, percent_elim = NULL)


rfun_EliminationTime <- function(out){
  df <- out %>% 
    select(starts_with("I"), time, sim)%>% 
    tidyr::pivot_longer(cols=!c(time, sim), names_to="patch")%>%
    filter(value != 0)%>%
    group_by(sim, patch)%>%
    summarize(last_case= max(time))%>%
    ungroup()%>%
    mutate(elim_patch = ifelse(last_case == 1095, 1, 0))
  
  df.elim <- df %>% #only for patches that do get eliminated
    group_by(sim)%>%
    summarise(elim_system = sum(elim_patch))%>%
    filter(elim_system == 0)
  
  num_elim = nrow(df.elim)
    
  
  df <- df %>%
    filter(elim_patch == 0)%>%
    group_by(patch)%>%
    summarise(mean = round(mean(last_case),0),
              SI_2.5 = round(quantile(last_case, .025),0),
              SI_97.5 = round(quantile(last_case, .975),0),
              num_sim=n())
  
  df.system_elimination <<- rbind(df.system_elimination, c(max(df$mean), num_elim/sim))

  
  return(knitr::kable(df))
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 2: Staggered pulse, all vaccinated (patch 1 --\> patch 2 --\> patch 3)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 3: Staggered pulse, all vaccinated (patch 2 --\> patch 1 --\> patch 3)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)


```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 4: Instant pulse, 2 vaccinated (70%), 1 (patch 1) vaccinated (30%)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 5: Instant pulse, 2 vaccinated (70%), 1 (patch 2) vaccinated (30%)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 6: Staggered pulse, all vaccinated (patch 1 (70%) --\> patch 2 (30%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 7: Staggered pulse, all vaccinated (patch 2 (70%) --\> patch 1 (30%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 8: Staggered pulse, all vaccinated (patch 1 (30%) --\> patch 2 (70%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 9: Staggered pulse, all vaccinated (patch 2 (30%) --\> patch 1 (70%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

## PART B: 3 patches: patch 1 \<-\> patch 2 \<-\> patch 3

```{r}
#contact matrix
df.contact <-data.frame(patch1 = c(0,1,0), patch2= c(1,0,1), patch3 =c(0,1,0))
beta.contact_constant = 0.00001 
beta.contact = rfun.FormatContact(df.contact, beta.contact_constant)

```

Sim 1: Single pulse, all vaccinated

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)),] = c(rep(0.7, num_patch)) #

#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 2: Staggered pulse, all vaccinated (patch 1 --\> patch 2 --\> patch 3)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 3: Staggered pulse, all vaccinated (patch 1 --\> patch 3 --\> patch 2)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 3] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 2] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 4: Staggered pulse, all vaccinated (patch 2 --\> patch 1 --\> patch 3)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 5: Instant pulse, 2 vaccinated (70%), 1 (patch 1) vaccinated (30%)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 6: Instant pulse, 2 vaccinated (70%), 1 (patch 2) vaccinated (30%)

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 7: Staggered pulse, all vaccinated (patch 1 (70%) --\> patch 2 (30%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 8: Staggered pulse, all vaccinated (patch 2 (70%) --\> patch 1 (30%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 9: Staggered pulse, all vaccinated (patch 1 (30%) --\> patch 2 (70%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 10: Staggered pulse, all vaccinated (patch 2 (30%) --\> patch 1 (70%) --\> patch 3 (70%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.3, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.7, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 11: Staggered pulse, all vaccinated (patch 1 (70%) --\> patch 2 (70%) --\> patch 3 (30%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.3, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```

Sim 12: Staggered pulse, all vaccinated (patch 2 (70%) --\> patch 1 (70%) --\> patch 3 (30%))

```{r, warning=FALSE, message=FALSE}
set.seed(123)

#set up VANCAN pulse schedules
vax <- vax0
vax[c(50, (50+365), (50+2*365)), 2] = c(rep(0.7, num_patch)) #
vax[c(100, (100+365), (100+2*365)), 1] = c(rep(0.7, num_patch)) #
vax[c(150, (150+365), (150+2*365)), 3] = c(rep(0.3, num_patch)) #


#save simulations together in a big df
out <- NULL
for(i in 1:sim){
  sub <-  rfun.PatchSEIV(init, parms, beta.contact, vax, Time, dt)%>%
    mutate(sim = i)
  out <- rbind(out, sub)
}

rfun_RabiesDynamics(out, parms$patch)
rfun_EliminationTime(out)

```

```{r, echo=FALSE, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
ggarrange(plotlist=fig1, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)
ggarrange(plotlist=fig2, ncol=3, nrow=ceiling(num_patch/2), common.legend = TRUE)

```

```{r, fig.width=8, fig.height=4, echo=FALSE, warning=FALSE, message=FALSE}
fig3
```


Compare systems


```{r, warning=FALSE, message=FALSE}
colnames(df.system_elimination) = c("Avg_time_elim", "Percent_elim")

df.sys <- df.system_elimination %>%
  mutate(Scenario = seq(1:length(Percent_elim)))%>%
  select(Scenario, Percent_elim, Avg_time_elim)


knitr::kable(df.sys)
```